Weld pool detection and geometrical analysis during GTAW process

### 2011-06-20 articles/TIG/paper_bain.tex

\title{ Weld pool detection and geometrical analysis during GTAW process by image processing library}

@@

Gas Tungsten Arc Welding (GTAW), which uses a non-consumable tungsten 
electrode and an inert gas for arc shielding, is an extremely 
important arc welding process. It is commonly used for welding hard-to-weld
 metals such as stainless steel \cite{JUANG} and therefore widely use in 
modern and basic industries. To achieve good weld quality in GTAW process,
 several weldment characteristics or objects should be sensed and
 controlled \cite{DOUMANIDIS}. It is shown that the geometry of the weld 
pool, in particular his shape and size, contains sufficient information
 about weld penetration to evaluate the weld quality \cite{ZHANG}.
 Furthermore it has been shown that the welding arc parameters to 
simulation and weld quality are directly relate to the weld pool 
geometry \cite{LU}.
Basically the weld pool object is the key in quality control 
to automated welding process \cite{KOVACEVIC}. 
According to this a better comprehension of weld pool behavior
 presents a GTAW process, could help to improve numerical 
simulations and enhance welding quality in manufacturing 
process \cite{LIN}, \cite{WU1}.

There have been many studies on visual sensing techniques
 for observing weld pool image \cite{BAE}. Optical sensors
 like high speed CCD cameras and lighting systems have been
 widely use in GTAW process to realize image acquisition 
\cite{GUANGJUN}, control process \cite{BAE} and parametric 
studies \cite{BALSAMO}. Then image processing plays a critical
 role in extracting useful information from visual scenes 
\cite{WANG}. Nevertheless the strong interference from the
 arc lightning required more than standard image treatment
 to analyses the raw images of the welding process 
\cite{NORDBRUCH}. Previous work has shown that is possible
 to perform geometrical analysis in weld pools \cite{WU1}.
 Parameters such as weld pool contour and surface has been
 detected and measure using different and specifics processing 
images algorithms \cite{KOVACEVIC}, \cite{WU1}, \cite{SAEED}.
 However, to date, effectively automatic images processing 
of GTAW process has not been developed, possibly due to the
 level difficulty involved for welding researchers \cite{WANG}.

To perform geometrical analysis in weld pool; a multipurpose
 C++ based library (erCv) was developed by the Weld/Assembly
 group of LMGC laboratory. These library results from
 selected functions join from highly reliable open source
 libraries to images treatment, geometrical analysis, 
graph theory applications and image visualization.

In a first step, and despite the different static and dynamic 
weld conditions, as well as different current regime, a 
reliable 2D profiles and geometrical parameters from weld
 pool, has been obtained using the mentioned library. 
Using this, a dimension and dimensionless analysis has 
been performed into a static and dynamic GTAW process. 
The intention is contribute to simplify futures numerical
 models of weld pool-arc interaction and quality process.

@@

Some definitions and brief description of generals principles used in image processing are required to better understand how is work the algorithms utilized by the erCv library.

@@

A numerical grey image can be describe as 3D surface discretized by a grid 
mesh in the X,Y plane. Each mesh represents a pixel and the relief surface
 at Z axis, the grey level.

A strong relief change or high gradient greyness values at the image are
 perceive by human eyes as light changes, and can be interpreted as 
objects edges. 
Sometimes, a regular relief patrons or regular greyness variation 
can be distinguish. The human eyes can perceive these patrons as 
texture and interpret the space between different textures zones as edges.
There exist a large spectrum of algorithms to image treatment, 
in particular to edges detect. Most of them can be classified 
by the way that his operate above the image pixels and grey level.

@@

The filters are algorithms that operate as mathematical functions 
$f$ above the $X$, $Y$ or both axis of the image ($Z = f(X,Y)$,  
$f(X)$ or $f(Y)$)  modifying his greyness value or $Z$ component. 
Different kind of filters can be mentioned as median, Gaussian, 
impulse, adaptive and others. The impulse filters such Canny are
 widely use to edge detection \cite{COCQUEREZ}. These filter have
 an impulse response to most important greyness gradient in the 
image; this allows the filter a better edges localisation (see 
figure \ref{photo-explication-filter}). For this reason, Canny 
filter is widely use at erCv library to detect the welding 
elements edges. However, it is sensible to noise or secondary
 greyness gradients in the image and, in consequence, it have 
some difficulties to define closer surface.

@@

\section{Image treatment library (erCv)}
\label{image-treatment-library-ercv}

As mentioned, a multipurpose image processing library was developed
 and currently use at the laboratory, in order to analyze welding
 process objects.
erCv is able to perform edge detection and geometrical analysis
 in a different welding objects such: Macro drop, droplets 
(for GMAW process) and weld pool (for GTAW process).
erCv is a modular library assembled in a oriented object C++ 
language. Then erCv is a scalable and portable library able 
to perform real time contour detection. erCv is build in Phyton 
scripts link to C++, making it relatively convivial to use.
erCv is compose by four processing modules based in C,  
C++ and python language (see figure \ref{schema-erCv}). These modules are: 
\begin{description}
\item[Image Treatment:] Due to weld process conditions such
 arc lightening, heat and electrodes positions; the raw image
 registered by CCD camera are not calibrated and present light
 inhomogeneities and noise. In order to obtain the real shape
 and size of weld elements, this module include calibrations
 algorithms. To detect the welding objects contours it is necessary
 to improve the weld element image. This module has the pre-processing
 treatments to noise reduction and image enhancement. Then to start
 the edges detection process, this module includes processing 
algorithms as segmentations by samples comparators, watershed
 transformation, filters edge detectors and histogram based
 methods.  
\item[Geometrical Treatment and Analysis:] This module has 
to end the edge detection process, completing and in some 
cases extrapolating the weld elements edge. It is also 
responsible to compute the geometrical data of welding 
elements such weld pool surface and metal transfer drop 
volumes. This module uses a full geometry algorithm library,
 which include different algorithms such as triangulations 
and mesh generation, alpha shape and convex hull generation
 and polygonal structures.
\item[Graph Theories:] To compute the geometrical data of
 welding elements it is necessary to extract the welding 
object edge from the image; this required some criteria 
such as continuity, length or closer condition.  This module 
use graph algorithms to convert edges pixels points into 
connected segments, and therefore, identify and select the 
welding element edge using the criteria. This module is compose
 by connect segments, estimates minimal cut, determine largest
 chain segments and others algorithms. 
\item[Visualization:] This module is a set of functions use
 to execute, show and/or register the different steps at the
 image process.
\end{description}

@@

To perform the geometrical analyses, different signals have to be 
synchronizes and recorder \cite{CHAPUIS}; therefore an accurate,
 reliable and synchronizes systems are requires due to the high
 amount of data and highly noisy environment (electromagnetic 
noise and arc light radiation).

A platform has been developed at the laboratory to perform multi-physics
 measures in arc welding process (see figure \ref{schema-platform}). 
The platform was conceive with an automatically procedure to synchronize,
 to acquire, to manage and to exploit large flow of multi-physical
 experimental data (up to 2 Go per test). This characteristic allows
 synchronizing (in time) the current and voltage signals with 
the acquired images.

@@

Stationary spots weld are made using the GMAW process with the Oerlikon
 CitoWave 500 generator. The target is a steel disk of $10\ mm$ of
 thickness and ER70S steel welding wire.
The test campaign began by a reference test, with parameters fixed
 to:  welding time $4\ s$, wire feed speed $6\ m/min$, frequency 
droplets $113\ Hz$ and percentage of shielding gas ($CO_{2}$) $8$
 (so $92\%$ argon). Welding parameters values are summarize in 
tables \ref{table-parameters-static} to static and
 \ref{table-parameters-change} to variables parameters.

\begin{table}
\begin{center}
\begin{tabular}{|cc|}
\hline
Welding wire type & ER70S \\ 
Wire diameter ($mm$) or $drw$ & 1 \\  
Contact tip to work distance ($mm$) & 20 \\
Shielding gaz flowrate ($l/min$) & 18 \\ \hline
\end{tabular}
\caption{{\small Determined constant welding parameters used in experiments}}
\label{table-parameters-static}
\end{center}
\end{table}

\begin{table}
begin{center}
\begin{tabular}{|cc|}
\hline
Welding wire type & ER70S \\ 
Wire diameter ($mm$) or $drw$ & 1 \\  
Contact tip to work distance ($mm$) & 20 \\
Shielding gaz flowrate ($l/min$) & 18 \\ \hline
\end{tabular}
\caption{{\small Variable welding parameters used in experiments}}
\label{table-parameters-change}
\end{center}
\end{table}

Welding current and arc voltage are recorded at $30\ kHz$ 
sampling rate. The images are recorded at $4000$ frames per
 seconds, which is enough to measure macro drop radius and 
apparent liquid-solid contact angle histories. The images 
and electrical signals are synchronize, thanks to the 
automatically approach made in the multi physics platform.

@@

The purpose is to study the evolution of weld pool 
object in a GTAW process. Therefore it is necessary 
to detect the contour, without predetermined shape, 
of the weld pool and calculate his surface area.

@@

. Most of them required complex image analysis using
 hierarchy decomposition (structural), fractal or stochastic model
s (model based) and Fourier or wavelet transforms (transforms methods) 
 Di BLA BLA use this  has been thedescribe  where the shadows zones are the projection of the macro drop and droplet.

>>>>

\title{ Weld pool detection and geometrical analysis during GTAW process}

\author{E Romero, J Chapuis ,C Bordreuil, F Souli\'e, G Fras}

\address{Laboratoire de Mécanique et Génie Civil,
CC048, Place Eugène Bataillon, Université Montpellier 2,
34095 Montpellier, France}
\ead{cyril.bordreuil@univ-montp2.fr}

@@

Gas Tungsten Arc Welding (GTAW), which uses a non-consumable tungsten 
electrode and an inert gas for arc shielding, is an extremely 
important arc welding process. It is commonly used for welding hard-to-weld
 metals such as stainless steel \cite{JUANG} and therefore widely use in 
modern and basic industries. To achieve good weld quality in GTAW process,
 several weldment characteristics or objects should be sensed and
 controlled \cite{DOUMANIDIS}. It is shown that the geometry of the weld 
pool, in particular its shape and size, contains sufficient information
 about weld penetration to evaluate the weld quality \cite{ZHANG}.
 Furthermore it has been shown that the welding arc parameters to 
simulation and weld quality are directly relate to the weld pool 
geometry \cite{LU}.
Basically the weld pool object is the key in quality control 
to automated welding process \cite{KOVACEVIC}. 
According to this a better comprehension of weld pool behavior
 presents a GTAW process, could help to improve numerical 
simulations and enhance welding quality in manufacturing 
process \cite{LIN}, \cite{WU1}.

There have been many studies on visual sensing techniques
 for observing weld pool image \cite{BAE}. Optical sensors
 like high speed CCD cameras and lighting systems have been
 widely use in GTAW process to realize image acquisition 
\cite{GUANGJUN}, control process \cite{BAE} and parametric 
studies \cite{BALSAMO}. Then image processing plays a critical
 role in extracting useful information from visual scenes 
\cite{WANG}. Nevertheless, the strong interference from the
 arc lightning required more than standard image treatment
 to analyses the raw images of the welding process 
\cite{NORDBRUCH}. Previous work has shown that is possible
 to perform geometrical analysis in weld pools \cite{WU1}.
 Parameters such as weld pool contour and surface has been
 detected and measure using different and specifics processing 
images algorithms \cite{KOVACEVIC}, \cite{WU1}, \cite{SAEED}.
 %However, to date, effectively automatic images processing 
%of GTAW process has not been developed, possibly due to the
% level difficulty involved for welding researchers \cite{WANG}.

To perform geometrical analysis in weld pool in real time monitoring, a multipurpose
 C++ based library (erCv) was developed at the lab. 
The purpose of the library is to monitor the weld pool and to extract geometrical 
information as weld pool surface or contour.
This library results from
 selected functions join from highly reliable open source
 libraries to images treatment, geometrical analysis, 
graph theory applications and image visualization.
The originality of the implemented library comes from 
the aggregation of the different libraries, each contributing to solve different
issues.

In a first step, and despite the different static and dynamic 
weld conditions, as well as different current regime, a 
reliable 2D profiles and geometrical parameters from weld
 pool, has been obtained using the mentioned library. 
Using this, a dimension and dimensionless analysis has 
been performed into a static and dynamic GTAW process. 
The intention is to contribute to enhance futur numerical
 models of weld pool-arc interaction and quality process.

@@

Some definitions and brief description of generals principles used in image 
processing are required to better understand how the algorithms are gathered by the erCv library.

@@

A numerical grey image can be described as 3D surface discretized by a grid 
mesh in the X,Y plane. Each mesh represents a pixel and the relief surface
 at Z axis, the grey level.

A strong relief change or high gradient greyness values at the image are
 perceive by human eyes as light changes, and can be interpreted as 
objects edges. 
Sometimes, a regular relief patrons or regular greyness variation 
can be distinguish. The human eyes can perceive these patrons as 
texture and interpret the space between different textures zones as edges.
There exist a large spectrum of algorithms to image treatment, 
in particular to edges detection. Most of them can be classified 
by the way that is operated on the image pixels and grey level.

@@

The filters are algorithms that operate as mathematical functions 
$f$ above the $X$, $Y$ or both axis of the image ($Z = f(X,Y)$,  
$f(X)$ or $f(Y)$)  modifying his greyness value or $Z$ component. 
Different kind of filters can be mentioned as median, Gaussian, 
impulse, adaptive and others. The impulse filters such Canny are
 widely use for edge detection \cite{COCQUEREZ}. These filter have
 an impulse response to most important greyness gradient in the 
image; this allows the filter a better edges localisation (see 
figure \ref{photo-explication-filter}). For this reason, Canny 
filter is widely use at erCv library to detect the welding 
elements edges. However, it is sensible to noise or secondary
 greyness gradients in the image and, in consequence, it have 
some difficulties to define closer surface.

@@

\section{Image treatment library}
\label{image-treatment-library-ercv}

As mentioned, a multipurpose image processing library was developed
 and currently use at the laboratory, in order to analyze welding
 process objects.
erCv is able to perform edge detection and geometrical analysis
 in a different welding objects such: Macro drop, droplets 
(for GMAW process) \cite{ROMERO1} and weld pool (for GTAW process).
erCv is a modular library assembled in a oriented object C++ 
language. The library is a scalable and portable library able 
to perform real time contour detection. erCv is build in Phyton 
scripts link to C++, making it relatively convivial to use.
erCv is composed by four processing modules based in C,  
C++ and python language (see figure \ref{schema-erCv}). These modules are: 
\begin{description}
\item[Image Treatment:] Due to weld process conditions such
 arc lightening, heat and electrodes positions; the raw image
 registered by CCD camera are not calibrated and present light
 inhomogeneities and noise. In order to obtain the real shape
 and size of weld elements, this module include calibrations
 algorithms. To detect the welding object contours: it is necessary
 to improve the weld element image. This module has the pre-processing
 treatments to noise reduction and image enhancement. Then to start
 the edges detection process, this module includes processing 
 algorithms as segmentations by samples comparators, watershed
 transformation, filters edge detectors and histogram based
 methods.  Basic algorithms come from OpenCv \cite{OPENCV} and are enriched 
 by some implementations.
\item[Geometrical Treatment and Analysis:] This module has 
to close the edge detection process, completing and in some 
cases extrapolating the weld elements edge. It is also 
responsible to compute the geometrical data of welding 
elements such weld pool surface and metal transfer drop 
volumes. This module uses a full geometry algorithm library,
 which include different algorithms such as triangulations 
and mesh generation, alpha shape and convex hull generation
 and polygonal structures. Basic algorithm come from CGAL \cite{CGAL} 
\item[Graph Theories:] To compute the geometrical data of
 welding elements it is necessary to extract the welding 
object edge from the image; this required some criteria 
such as continuity, length or closure condition.  This module 
use graph algorithms to convert edges pixels points into 
connected segments, and therefore, identify and select the 
welding element edge using the criteria. This module is compose
 by connect segments, estimates minimal cut, determine largest
 chain segments and others algorithms. 
\item[Visualization:] This module is a set of functions use
 to execute, show and/or register the different steps at the
 image process.
\end{description}

@@

To perform the geometrical analysis, different signals have to be 
synchronized and recorded \cite{CHAPUIS}; therefore an accurate,
 reliable and synchronized systems are required due to the high
 amount of data and highly noisy environment (electromagnetic 
noise and arc light radiation).

A platform has been developed at the laboratory to perform multi-physics
 measures in arc welding process (see figure \ref{schema-platform}). 
The platform was designed with an automatically procedure to synchronize,
 to acquire, to manage and to exploit large flow of multi-physical
 experimental data (up to 2 Go per test). This characteristic allows
 synchronizing (in time) the current and voltage signals with 
the acquired images.

@@

Stationary spots weld are made using the GTAW process with the Sincosald
 generator. The target is a steel disk of $10\ mm$ of
 thickness and ER70S steel welding wire.
The test campaign began by a reference test, with parameters fixed
 to:  welding time $4\ s$ and the shielding gas is 
 $100\%$ argon. Welding parameters values are summarized in 
tables \ref{table-parameters-static} to static and
 \ref{table-parameters-change} to variables parameters.

For weld bead on a flat plate, a travel speed of 30cm/min is chosen.
\begin{table}
\begin{center}
\begin{tabular}{|cc|}
\hline
Welding wire type & ER70S \\ 
Contact tip to work distance ($mm$) & 20 \\
Shielding gaz flowrate ($l/min$) & 18 \\ \hline
\end{tabular}
\caption{{\small Determined constant welding parameters used in experiments}}
\label{table-parameters-static}
\end{center}
\end{table}

\begin{table}
begin{center}
\begin{tabular}{|cc|}
\hline
Welding wire type & ER70S \\ 
Wire diameter ($mm$) or $drw$ & 1 \\  
Contact tip to work distance ($mm$) & 20 \\
Travel Speed (cm/min) & 30 \\
Shielding gaz flowrate ($l/min$) & 18 \\ \hline
\end{tabular}
\caption{{\small Variable welding parameters used in experiments}}
\label{table-parameters-change}
\end{center}
\end{table}

Welding current and arc voltage are recorded at $30\ kHz$ 
sampling rate. The images are recorded at $1000$ frames per
 seconds, which is enough to measure liquid-solid interface. The images 
and electrical signals are synchronize, thanks to the 
automatically approach made in the multi physics platform.

@@

The purpose is to study the evolution of weld pool 
object in a GTAW process. Therefore it is necessary 
to detect the contour (polygon), without predetermined shape, 
of the weld pool and calculate its surface area.

@@

Most of them required complex image analysis using
 hierarchy decomposition (structural), fractal or stochastic model
s (model based) and Fourier or wavelet transforms (transforms methods) 
 Di BLA BLA use this  has been thedescribe  where the shadows zones are the projection of the macro drop and droplet.
